---
title: "Supplementary Material: Worse Prognosis, Larger Benefit"
subtitle: "A Bayesian Hierarchical Ordinal Meta-analysis of 30 Thrombectomy Trials"
author:
  - name: Bradley Kolb, MD
    affiliation: Department of Neurosurgery, Rush University Medical Center
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: false
    theme: cosmo
    embed-resources: true
  pdf:
    toc: true
    number-sections: false
    geometry: margin=1in
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{pdflscape}
        \usepackage[table]{xcolor}
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: png
    dpi: 300
bibliography: ../Bibliography_base.bib
csl: vancouver.csl
nocite: |
  @Ciccone2013_synthesis, @Broderick2013_ims3, @Kidwell2013_mrrescue,
  @Berkhemer2015_mrclean, @Campbell2015_extendia, @Goyal2015_escape,
  @Saver2015_swiftprime, @Jovin2015_revascat,
  @Bendszus2016_thrill, @Mocco2016_therapy, @Bracard2016_thrace,
  @Khoury2017_easi, @Muir2017_piste,
  @Albers2018_defuse3, @Nogueira2018_dawn,
  @Liu2020_best, @Martins2020_resilient,
  @Langezaal2021_basics,
  @Mocco2022_positive, @Tao2022_attention, @Jovin2022_baoche,
  @Yoshimura2022_rescuejapan,
  @Bendszus2023_tension, @Olthuis2023_mrclean_late,
  @Huo2023_angel, @Sarraj2023_select2,
  @Costalat2024_laste, @Yoo2024_tesla,
  @Goyal2025_mvo, @Psychogios2025_distal
---

```{=latex}
\clearpage
```

## Search Strategy

We searched PubMed (MEDLINE via NCBI) from inception to October 19, 2025, with no date limits. The full, reproducible PubMed strategy was:

```
(
"mechanical thrombectomy"[tiab] OR "endovascular thrombectomy"[tiab] OR
"aspiration thrombectomy"[tiab] OR "stent retriever*"[tiab] OR stentriever*[tiab] OR
Solitaire[tiab] OR Trevo[tiab] OR Penumbra[tiab]
)
AND
(
stroke[tiab] OR "ischemic stroke"[tiab] OR "large vessel occlusion"[tiab] OR LVO[tiab] OR
"basilar artery"[tiab] OR "basilar occlusion"[tiab]
)
AND
(
randomized controlled trial[Publication Type] OR random*[tiab] OR trial[tiab]
)
NOT (animals[mh] NOT humans[mh])
AND english[lang]
```

This search returned 327 records.

```{=latex}
\clearpage
```

## Eligibility Criteria

- **Design:** Randomized.
- **Population:** Adults with acute ischemic stroke (anterior or posterior circulation).
- **Intervention:** Mechanical thrombectomy (including stent retrievers and/or aspiration).
- **Comparator:** Best medical management (with or without IV thrombolysis).
- **Outcomes:** Functional outcomes (modified Rankin Scale).
- **Exclusions:** Non-randomized studies.

```{=latex}
\clearpage
```

## Study Selection and Data Handling

We screened titles/abstracts and then full texts, with disagreements resolved by consensus, as documented in the PRISMA 2020 diagram below.

For each included trial we extracted trial design, population, time window, imaging selection, intervention details (device/approach), comparator, and outcomes (including full-ordinal mRS where available).

For mRS results that were presented as percentages, we obtained mRS counts by assuming the reported percentage of patients achieving the given mRS score was obtained by dividing the number of patients achieving that score in the intention-to-treat population by the total number of patients in the intention-to-treat population across all 6 mRS categories.

```{=latex}
\clearpage
```

## PRISMA Diagram

![PRISMA 2020 diagram of search strategy and study selection.](../Figures/prisma.svg){width=100%}

```{=latex}
\clearpage
```

## Detailed References

All 30 included trials, listed in chronological order of publication.

::: {#refs}
:::

```{=latex}
\clearpage
\begin{landscape}
```

## Trial Characteristics

```{r}
#| label: tbl-s1

library(dplyr)
library(knitr)
library(kableExtra)

trial_meta <- read.csv("../master_supporting_docs/supporting_code/trial_metadata.csv",
                       check.names = FALSE, stringsAsFactors = FALSE)

trial_meta |>
  arrange(year) |>
  select(trial_display, year, n_evt, n_control, occlusion, time_window, population,
         primary_endpoint, key_feature, notes) |>
  kable(
    booktabs  = TRUE,
    longtable = TRUE,
    linesep   = "",
    col.names = c("Trial", "Year", "N (EVT)", "N (Ctrl)", "Occlusion",
                  "Time Window", "Population", "Primary Endpoint",
                  "Key Feature", "Notes")
  ) |>
  kable_styling(
    latex_options = c("repeat_header"),
    font_size = 8
  ) |>
  row_spec(0, bold = TRUE, background = "#f0f0f0") |>
  column_spec(1,  width = "2.0cm") |>
  column_spec(2,  width = "0.8cm") |>
  column_spec(3,  width = "0.9cm") |>
  column_spec(4,  width = "1.1cm") |>
  column_spec(5,  width = "1.7cm") |>
  column_spec(6,  width = "1.5cm") |>
  column_spec(7,  width = "3.0cm") |>
  column_spec(8,  width = "2.4cm") |>
  column_spec(9,  width = "2.8cm") |>
  column_spec(10, width = "1.5cm")
```

```{=latex}
\end{landscape}
\clearpage
\begin{landscape}
```

## Risk of Bias Assessment

```{r}
#| label: tbl-s2

library(dplyr)
library(knitr)
library(kableExtra)

rob <- read.csv("../master_supporting_docs/supporting_code/risk_of_bias.csv",
                check.names = FALSE, stringsAsFactors = FALSE)

domain_cols <- c("Random Sequence", "Allocation Concealment",
                 "Blinding (Performance)", "Blinding (Detection)",
                 "Attrition", "Selective Reporting", "Other")

rob_display <- rob |>
  rename(
    Trial                    = trial_display,
    `Random Sequence`        = random_sequence,
    `Allocation Concealment` = allocation_concealment,
    `Blinding (Performance)` = blinding_performance,
    `Blinding (Detection)`   = blinding_detection,
    `Attrition`              = attrition,
    `Selective Reporting`    = selective_reporting,
    `Other`                  = other
  ) |>
  select(-trial)

get_colors <- function(vals) {
  dplyr::case_when(
    vals == "Low risk"     ~ "#d4edda",
    vals == "High risk"    ~ "#f8d7da",
    vals == "Unclear risk" ~ "#fff3cd",
    TRUE                   ~ "#ffffff"
  )
}

tbl <- rob_display |>
  kable(
    booktabs  = TRUE,
    longtable = TRUE,
    linesep   = ""
  ) |>
  kable_styling(
    latex_options = c("repeat_header"),
    font_size = 8
  ) |>
  row_spec(0, bold = TRUE, background = "#f0f0f0") |>
  column_spec(1, width = "2.0cm") |>
  column_spec(2, width = "2.2cm") |>
  column_spec(3, width = "2.5cm") |>
  column_spec(4, width = "2.3cm") |>
  column_spec(5, width = "2.3cm") |>
  column_spec(6, width = "2.0cm") |>
  column_spec(7, width = "2.3cm") |>
  column_spec(8, width = "1.5cm")

for (col in domain_cols) {
  col_idx <- which(names(rob_display) == col)
  tbl <- column_spec(tbl, col_idx, background = get_colors(rob_display[[col]]))
}

tbl
```

```{=latex}
\end{landscape}
```

## Main Model Specification

```{r}
#| label: load-models
library(brms)
library(posterior)
options(width = 80)
fit_ordinal  <- readRDS("../master_supporting_docs/supporting_code/artifacts/fit_ordinal.rds")
fit_uv       <- readRDS("../master_supporting_docs/supporting_code/artifacts/fit_uv.rds")
fit_adjacent <- readRDS("../master_supporting_docs/supporting_code/artifacts/fit_adjacent.rds")
fit_bin      <- readRDS("../master_supporting_docs/supporting_code/artifacts/fit_bin.rds")
```

The main model was a Bayesian hierarchical proportional-odds cumulative ordinal model fit with `brms` (Stan backend via `cmdstanr`). The model formula was:

```{r}
#| label: main-model-output
print(fit_ordinal)
```

All Rhat values $\leq$ `r round(max(rhat(fit_ordinal), na.rm = TRUE), 2)`, consistent with convergence. Bulk and tail effective sample sizes were adequate (all >400 for key parameters).

```{=latex}
\clearpage
```

## Alternative Models

Three additional models were fit to assess robustness of the main findings.

### Alternative Model 1: Ordinal (PO + Unequal Variances)

Formula: `mrs_better | weights(count) ~ treatment + (1 + treatment | trial); disc ~ 0 + treatment`
Family: cumulative (logit link, log disc link)
Chains: 4 | Iterations: 2000 (1000 warmup) | Total post-warmup draws: 4000

```{r}
#| label: model-uv-output
print(fit_uv)
```

### Alternative Model 2: Ordinal (Adjacent-Category)

Formula: `mrs_better | weights(count) ~ treatment + (1 + treatment | trial)`
Family: acat (adjacent-category logit)
Chains: 4 | Iterations: 2000 (1000 warmup) | Total post-warmup draws: 4000

```{r}
#| label: model-adjacent-output
print(fit_adjacent)
```

### Alternative Model 3: Binary (mRS 0–2 vs 3–6)

Formula: `good_n | trials(total_n) ~ treatment + (1 + treatment | trial)`
Family: binomial (logit link)
Chains: 4 | Iterations: 2000 (1000 warmup) | Total post-warmup draws: 4000

```{r}
#| label: model-bin-output
print(fit_bin)
```

All three alternative models yield a consistently negative intercept–slope correlation, confirming robustness of the main finding.

## Supplementary Figure 1: Robustness of $\rho$ to Model Specification

Posterior medians and 95% credible intervals for $\rho$ under the primary proportional-odds ordinal model and three alternative models: an unequal-variance proportional-odds model, an adjacent-category ordinal model, and a binary model (mRS 0–2 vs 3–6). Estimates remain negative across all specifications.

![Robustness of the intercept–slope correlation to model specification.](../Figures/figure_three.svg){width=100%}

```{=latex}
\clearpage
```

## Supplementary Figure 2: Absolute Benefit Across Model Specifications

For each model, points show each trial's posterior median control-group probability of functional independence (x-axis) and posterior median absolute benefit in functional independence due to thrombectomy (y-axis). Functional independence is defined as mRS 0–2. The dashed line indicates no absolute benefit; the smooth curve is a descriptive fit with 95% uncertainty band.

![Baseline prognosis and absolute benefit in functional independence.](../Figures/figure_four.svg){width=100%}

```{=latex}
\clearpage
```

## Model Diagnostics

The following diagnostic figures are generated by running `run.R` and are saved to `master_supporting_docs/supporting_code/artifacts/`.

### Trace Plots

Markov chain trace plots for the six threshold parameters, the treatment effect, and the three variance-covariance hyperparameters.

```{r}
#| label: fig-trace
#| fig-cap: "Trace plots for main model parameters."
#| out-width: "100%"
#| eval: !expr file.exists("../master_supporting_docs/supporting_code/artifacts/ordinal_trace_plots.svg")

knitr::include_graphics("../master_supporting_docs/supporting_code/artifacts/ordinal_trace_plots.svg")
```

<!-- Developer note: If this figure is missing, run master_supporting_docs/supporting_code/run.R to generate diagnostic SVGs. -->

```{=latex}
\clearpage
```

### R-hat and Effective Sample Size

```{r}
#| label: fig-rhat
#| fig-cap: "R-hat values and effective sample sizes for main model."
#| out-width: "100%"
#| eval: !expr file.exists("../master_supporting_docs/supporting_code/artifacts/ordinal_convergence_values.svg")

knitr::include_graphics("../master_supporting_docs/supporting_code/artifacts/ordinal_convergence_values.svg")
```

<!-- Developer note: If this figure is missing, run master_supporting_docs/supporting_code/run.R to generate diagnostic SVGs. -->

```{=latex}
\clearpage
```

### Posterior Predictive Check

```{r}
#| label: fig-ppc
#| fig-cap: "Posterior predictive check for the main model, grouped by treatment arm."
#| out-width: "100%"
#| eval: !expr file.exists("../master_supporting_docs/supporting_code/artifacts/ordinal_posterior_predictive_check.svg")

knitr::include_graphics("../master_supporting_docs/supporting_code/artifacts/ordinal_posterior_predictive_check.svg")
```

<!-- Developer note: If this figure is missing, run master_supporting_docs/supporting_code/run.R to generate diagnostic SVGs. -->

```{=latex}
\clearpage
```

### Prior and Likelihood Sensitivity

Prior power-scaling sensitivity analysis for key model parameters. Sensitivity was low across all parameters.

```{r}
#| label: fig-prior
#| fig-cap: "Prior sensitivity (power-scaling) analysis for the main model."
#| out-width: "100%"
#| eval: !expr file.exists("../master_supporting_docs/supporting_code/artifacts/ordinal_power_scaling_sensitivity.svg")

knitr::include_graphics("../master_supporting_docs/supporting_code/artifacts/ordinal_power_scaling_sensitivity.svg")
```

<!-- Developer note: If this figure is missing, run master_supporting_docs/supporting_code/run.R to generate diagnostic SVGs. -->

```{=latex}
\clearpage
```